\section{Methods and Workplan}

The discussion of methods and workplan is organized according to
the two major objectives listed in the introduction.
Each objective is in turn elucidated in terms of three outcomes
(as listed at the conclusion of each of the preceding subsections).
The basic plan is that outcomes 1.1 and 2.1 will be pursued
during year 1; outcomes 1.2 and 2.2 during year 2;
and outcomes 1.3 and 2.3 during year 3.
The following discussion fleshes out the six outcomes in terms of the
tasks that must be performed in order to achieve them. The allocation of activities to project
personnel is discussed in the budget justification.


\subsection*{Objective 1: Access to Language Resources in Archives}

\emph{Develop guidelines and services that encourage best common
    practices among language archives that will 
    facilitate language resource discovery with precision through OLAC.}

\def\task{1.1}
\paragraph{Outcome {\task}: All OLAC repositories should have up-to-date catalogs
      that contain metadata conforming to best practice.}

\begin{enumerate}[label=\emph{\task\alph*}]\setlength{\itemsep}{0pt}
\item \emph{Best practice document:}
  Prepare and adopt an OLAC recommendation on metadata best practice,
  adapting the DC Usage Guidelines and fleshing out published recommendations
  \citep{BirdSimons03language,Simons06}.
\item \emph{Score cards:}
  Update the automated score card system that reports on the quality
  of metadata records to conform to the adopted recommendations.
\item \emph{Metadata quality:}
  Identify low-scoring archives and work with them to
  improve the quality of their metadata, particularly the
  use of OLAC vocabularies. 
  Develop an OLAC metadata usage note based on this experience.
\item \emph{Updating archives:}
  Identify OLAC archives whose metadata records have gone stale,
  and work with them to bring their catalogs up-to-date and to
  set up automatic processes to expose their live catalogs.
\item \emph{Quarterly reporting:}
  Develop an automated quarterly report to be emailed to curators to 
  remind them about the quality and currency of their metadata, and to
  inform them of statistics concerning queries on their archives.
\item \emph{Metrics:}
  Develop metrics for monitoring the size, coverage, and use of the 
  complete OLAC metadata catalog, and implement such reporting on the web site.

\end{enumerate}

\def\task{1.2}
\paragraph{Outcome {\task}: All major language archives should be participating in OLAC.}

\begin{enumerate}[label=\emph{\task\alph*}]\setlength{\itemsep}{0pt}
\item \emph{Non-participating archives:}
  Compile a list of all known non-participating language archives.
\item \emph{Collaboration:}
  Contact all identified archives and consult with them to develop best
  strategy for generating their metadata.
\item \emph{Scraping:}
  Do automatic scraping of the HTML metadata already published on
  archive sites and perform a quick low-accuracy conversion to OLAC
  metadata in order to show archivists their search results coming up
  in user queries; use this to show archivists the value of mapping their
  terms to OLAC terms and publishing OLAC metadata in order to
  improve precision and recall.
\item \emph{OLAC export:}
  Work with these archives to expose their catalog in OLAC format.
\item \emph{Repository editor:}
  Configure an open source XML editor (like XMLmind) to create a static repository editor
  suitable for use by small archives.
\end{enumerate}

\def\task{1.3}
\paragraph{Outcome {\task}: All OLAC repositories should conform to current best practices
      for the long-term curation of their holdings.}

\begin{enumerate}[label=\emph{\task\alph*}]\setlength{\itemsep}{0pt}
\item \emph{Recommendation:}
  Prepare and adopt an OLAC recommendation document on 
  best practices for the long-term curation of language archive holdings.
\item \emph{Archive categories:}
  Refine the {\tt <olac-archive>} description to include
  more fine-grained categorization of participating archives.
  Categories should be based on published criteria (in the OAIS reference model)
  and assignment of categories should be vetted by OLAC.
\item \emph{Score cards:}
  Incorporate these criteria into the automated score cards.
  Generalize the score card mechanism to facilitate the addition
  of new measures over time (as best practices evolve).
\end{enumerate}

\subsection*{Objective 2: Access to Language Resources on the Web}

\emph{Develop services to bridge the resource catalogs of the
  repository, library, and web domains (OAI, MARC, Google)
  to facilitate language resource discovery with precision through OLAC.}

\def\task{2.1}
\paragraph{Outcome {\task}:  All repository and library holdings relevant to language
   documentation should be indexed in OLAC, by
   crosswalking and enriching existing catalog records.}

\begin{enumerate}[label=\emph{\task\alph*}]\setlength{\itemsep}{0pt}
\item \emph{Language identification:}
  Develop an automated procedure for finding
  language identification in a catalog record and translating it to an
  appropriate ISO 639-3 code.
\item \emph{Data type identification:}
  Develop an automated procedure for finding
  linguistic data type information in a catalog record and translating
  it to an appropriate OLAC linguistic data type code.
\item \emph{OAI Crosswalk:}
  Develop an OLAC data provider that harvests from the
  leading OAI aggregator and serves up all records that are determined to
  be language resources with the metadata enriched for precise
  identification of language and linguistic data type (primary text vs
  lexicon vs linguistic description).
\item \emph{Z39.50 Crosswalk:}
  Develop an OLAC data provider that harvests from one or more
  Z39.50 gateways (e.g.~using JZKit), eliminates duplicates,
  applies the MARC to DC crosswalk\footnote{\scriptsize\url{http://www.loc.gov/marc/marc2dc.html}},
  and serves up all records that are determined to
  be language resources with the metadata enriched for precise
  identification of language and linguistic data type.
\end{enumerate}

% JZKit:
% Ian Ibbotson is the contact:
% http://www.loc.gov/z3950/agency/resources/software.html

\def\task{2.2}
\paragraph{Outcome {\task}: Low density language materials identified by
  linguistic web mining should be reliably categorized with OLAC vocabularies.}

\begin{enumerate}[label=\emph{\task\alph*}]\setlength{\itemsep}{0pt}
\item \emph{Language identification:}
  Develop an automated procedure for finding
  low-density language identification from an arbitrary web document,
  (e.g.~using character encodings, keywords,
  n-gram signatures, anchor text, and place-name references)
  and translating this to an appropriate ISO 639-3 code.
\item \emph{Data type identification:}
  Develop an automated procedure for identifying
  the linguistic data type of an arbitrary web document
  (e.g.~using Bayesian classifiers trained on existing metadata records)
  and translating this to an appropriate OLAC linguistic data type code.
\item \emph{Web mining:}
  Generate OLAC metadata records for low-density
  language resources discovered by linguistic web-mining projects,
  such as Hughes' collection of 600,000 low-density language URLs.
\end{enumerate}

% \item modify OLAC search to group results by item not by location

\def\task{2.3}
\paragraph{Outcome {\task}: Web search engines should index all OLAC records,
      so that users who discover language resources using a web search
      quickly find OLAC records and are drawn to the OLAC site for
      more precise searching.}

\begin{enumerate}[label=\emph{\task\alph*}]\setlength{\itemsep}{0pt}
\item \emph{Static HTML:}
  Generate a static HTML page for each OLAC metadata record, by adding
  a new module to the OLAC harvester.
\item \emph{Synonyms:}
  Add synonyms for all identified linguistic terminology to static HTML pages,
  so that these pages are more likely to be found in conventional web searches.
\item \emph{Embedded queries:}
  Enrich static pages with OLAC query links, so that users who discover
  OLAC records via a Google search result are encouraged to remain on the
  OLAC site for more precise searching.
\end{enumerate}

\section{Dissemination Plans}
\label{sec:dissemination}

We will present the results of our research at major international
conferences in linguistics, computational linguistics, and
digital libraries, and submit extended discussions of the research to
leading journals.  The results of our research will also be deployed
on the OLAC web site as they come into existence. In this way the
results will be put into action immediately by the institutions
who are contributing their resources to the OLAC catalog and by 
individuals who are using the OLAC search services to find them.
All tools and data created by the project will be
disseminated with open source and open content licenses.
We will make occasional use of LanguageLog, when appropriate, to
disseminate information of broader interest that is discovered using
OLAC search services.
