\section{Background and Results from Prior Support}

%%%%%%%% Scale...

Today, language technology and the linguistic sciences are confronted
with a vast array of \emph{language resources}.  These are
predominantly resources in or about a human language---texts, recordings, 
dictionaries, annotations, grammars, and the like.
Also included are resources used by language technologists---software,
protocols, data models, file formats, newsgroups, web
indexes, and the like.  The resources are growing
in number, in size, in diversity.  Multiple \emph{communities} depend
on language resources, including linguists, engineers, educators, students, and
actual speakers.  Language resources are particularly important to 
language communities who want to revitalize
a language that is in danger of extinction.
% Many individuals and institutions serve these communities,
% including archivists, software developers, and publishers.
The number of individuals having an immediate, professional interest in language
resources is in the tens of thousands. The total set of potential users
of language resources must number in the millions.

%%%%%%%% The opportunity...

Today, we have unprecedented opportunities to \emph{connect} these
communities to the language resources they need.  First, 
% inexpensive mass storage technology permits large resources
% to be stored in digital form,
the world wide web is making it possible, like never before, to bring
all the members of the language resource community together into 
a global virtual community. 
Second, digital publication---both on and off
the web---is the most practical and efficient means of
sharing language resources.  
Open encoding standards like the Extensible Markup Language (XML) 
and the Universal Character Set (Unicode)
provide flexible ways to represent structured data and ensure its
long-term survival. 
Finally, the Open Language Archives Community (OLAC) provides a 
standard set of language resource descriptors (OLAC Metadata), 
and a portal that permits users to query dozens of language 
archives simultaneously using a single search. OLAC has the
bottom-up, distributed character of the web, while simultaneously
having the efficient, structured nature of a centralized database.
This combination is well-suited to the language resource community.
% where the available data is growing rapidly and where a large
% user-base is consistent in how it describes its resource needs.

%%%%%%%% What OLAC has done...

OLAC was founded in 2000 with the goals of (i) developing consensus on
best current practice for the digital archiving of language resources,
and (ii) developing a network of interoperating repositories and
services for housing and accessing such resources.  For the first
goal, substantial progress on identifying best practices has now been
achieved under the NSF E-MELD project; consideration of best practices
is now incorporated into the funding of language documentation
projects (DEL, \citet{BirdSimons03language}).  For the second goal,
35 repositories holding some 30,000 metadata records are
participating in OLAC; linguists can search all repositories
simultaneously, and the interest level now stands
at 100,000 queries per month.  Participating archives include:
Alaska Native Language Center Archive,
Archive of the Indigenous Languages of Latin America,
UC Berkeley Audio Archive of Linguistic Fieldwork,
Oxford Text Archive, Rosetta Project All Language Archive, 
SIL Language and Culture Archives, and 29 more.

%%%%%%%% Shortcomings...

% Despite these early successes, it is still true today that ``new
% digital language resources of all kinds -- lexicons, interlinear
% texts, grammars, language maps, field notes, recordings -- are
% difficult to reuse and less portable than the conventional printed
% resources they replace'' \citep{BirdSimons03language}.  The proliferation of
% language ``archiving projects'' has only made this problem more
% acute: endangered languages are captured in ephemeral file formats,
% and the so-called ``archive'' goes away at the end of the project.
% Mirroring the twin goals of OLAC, there are just two things a linguist
% must do to ensure that digital data endure: ``(1) the materials must be
% put into an enduring file format, and (2) the materials must be
% deposited with an archive that will make a practice of migrating them
% to new storage media as needed'' (Simons 2006).  This amounts to an
% injunction to individuals to \emph{use} portable formats and credible
% archives.

The results to-date of OLAC demonstrate a high level of participation
and interest within the language resources community, but we recognize
that the current coverage is only the 
tip of the iceberg. Google (with billions of items indexed) is still
the first search engine of choice for most linguists since it has so
much greater coverage. However, using a conventional search engine to
find a particular kind of resource for a particular language is
not likely to succeed, especially for an uncommon language. 
One of three outcomes is more likely:
(1) Relevant resources do exist as web pages, but they are 
{\it occluded} from view by fuzziness of search terms and 
descriptors on the page.
(2) Relevant resources are accessible from the web, but they are
{\it hidden} from the view of the search engine because they are
on the other side of form-based interfaces.
(3) Relevant resources are {\it missing} altogether because the
archive has not published its catalog or because they have not
been archived by the linguists who created them.

% The proposed project will create new incentives that are designed to
% encourage linguists to adopt these new practices.  The project has
% three objectives:

The aim of the proposed project is to dramatically improve the language
resource community's access to resources by achieving an 
order-of-magnitude increase in the coverage of the OLAC catalog
and in the use of OLAC search services.
The project seeks to do so through three main areas of activity:

\begin{description}
  \item[Objective 1: Access to Language Resources in Archives:]
    Develop guidelines and services that encourage conformance to best
    practices in cataloging and curating the world's language resources
    in language archives.

  \item[Objective 2: Access to Language Resources on the Web:]
    Develop services to bridge the resource catalogs of the
    repository, library, and web domains,
    to facilitate language resource discovery with precision through OLAC.

  \item[Objective 3: Access to Language Resources from the Field:]
    Extend institutional repository software to support the
    best-practice cataloging and curation of language resources
    uploaded by linguists themselves.
\end{description}

For the most part, objective 1 addresses the problem of occluded resources,
objective 2 addresses hidden resources, and objective 3 addresses missing resources.
In the longer term, it is hoped that an OLAC search
service with vastly increased coverage will become such a popular tool
among linguists that it will provide powerful incentive for them to
adopt best practices in digital archiving of their own work so that it
can become accessible to others through this means. 

\textbf{Prior Support:} The proposed project represents an outgrowth of
prior NSF support to the investigators in the following projects:
\textit{Multidimensional Exploration of Linguistic Databases}
  ({\small \#9983258, 02/01/2000--01/31/2003});
\textit{ISLE: International Standards for Language Engineering}
  ({\small \#9910603, 03/01/2000--07/31/2003});
\textit{TalkBank: A Multimedia Database of Communicative Interactions}
  ({\small \#9978056, 10/01/1999--09/30/2005});
\textit{E-MELD: Electronic Metastructure for Endangered Languages Data}
  ({\small \#0094934, 07/01/2001--06/30/2006});
\textit{The Rosetta Project: ALL Language Archive}
  ({\small \#0333727, 11/01/2003--02/28/2006});
\textit{Querying Linguistic Databases}
  ({\small \#0317826, 08/01/2003--07/31/2007}).
Key results from these projects were as follows:

\renewcommand{\labelenumi}{(\alph{enumi})}
\begin{enumerate}\setlength{\itemsep}{0pt}
\item OLAC, a community-wide resource discovery framework
\citep{BirdSimons00,BirdSimons00survey,BirdSimons01,BirdSimons02workshop,Simons02query,BirdSimons03chum,SimonsBird03lht,Simons03display,SimonsBird03llc,BirdSimons04metadata};

\item data models for linguistic annotations of text and speech
\citep{BirdLiberman01,MaedaBird00,GraffBird00,CottonBird02,CieriBird01,ATLAS00,BirdHarrington01}

\item open source linguistic annotation software
\citep{Bird01acl,MaedaBird02,BirdMaeda02,MaLee02};

\item recommendations for digital storage of language data
\citep{BirdSimons03language,Simons06};

\item methods of interoperating over disparate language resources
\citep{SimonsEtal04text,SimonsEtal04lex};

\item CD-ROM publications of linguistic field data:
\citep{BirdBell01,Bird03ngomba,Bird03paradigms};
and

\item models and tools for querying linguistic databases
\citep{BirdBuneman01,BirdBunemanTan00,LaiBird04,Bird05planx,Bird06icde}
\end{enumerate}

\subsection{Access to Language Resources in Archives}

%%%%%%%% The status quo...

Members of the language resources community have a common need for
discovering language resources with high precision and recall.  To
support this, the community needs to agree on
some specialized vocabularies in the metadata for
describing resources.  Over the past five years, OLAC
participants have adopted five such vocabularies:
\textit{subject language},
  for identifying precisely which language(s) a resource is ``about'';
\textit{linguistic type},
  for classifying the structure of a resource as primary text,
  lexicon, or language description;
\textit{linguistic field},
  for specifying relevant subfields of linguistics;
\textit{discourse type},
  for indicating the linguistic genre of the material;
  and
\textit{role},
  for documenting the parts played by specific individuals and institutions
  in creating a resource.
These vocabularies are represented in the OLAC Metadata format,
which is an extension of Dublin Core Metadata \citep{BirdSimons04metadata}---the dominant metadata standard in the
digital library and world wide web communities.
Participating language archives publish their catalogs in an XML
format, and these records are ``harvested'' twice a day by OLAC services
using the Open Archives Initiative (OAI) Protocol for Metadata Harvesting
\citep{SimonsBird03lht}---another standard of the digital library community.

%%%%%%%% How OLAC has been innovative...

OLAC has been a developer or early adopter of several other key
digital library technologies in service to the language resources
community:
(i) a simplified method for archives to publish their metadata in
  ``static repositories'', thus lowering the barrier to entry---a
  method now adopted by the OAI as a service to the wider
  digital archives community;
(ii) a Google-like search interface for searching participating
  archives with specialised support for language identification
  \citep{HughesKamat05,Hughes06lrec};
(iii) a report card system which publishes a metadata quality score
  for each participating archive to serve as a form of
  peer review and an incentive for archives to improve the quality of
  their metadata;
(iv) a metadata usage report showing how individual metadata
  elements and values have been put to use by participating archives;
(v) an ``OAI crosswalk'' permitting users of digital library services
  outside the language resources community to access all OLAC metadata;
  and
(vi) a web crawler gateway permitting users of search engines such
  as Google to access all OLAC metadata.

Beyond these technological achievements, OLAC has succeeded in
establishing a \emph{community} that operates by means of an open
process \citep{OLAC-Process}.  In a 2004 publication of the
Digital Library Federation, OLAC was singled out for this
facet of its work:

\begin{quote} \small
  OLAC is exemplary in several ways: the technical and social
  infrastructure that it has developed to support its community of
  contributors, based on shared principles and standards; the
  resources that it provides at its Web site about its purpose, scope,
  history, tools, news and events; and the efforts of its two leaders
  ... to articulate the challenges, analyze the options, and recommend
  possible solutions to their community of contributors in order to
  improve OLAC \citep{Brogan04}.
\end{quote}

%%%%%%%% Three shortcomings...

Despite these successes, OLAC has three significant shortcomings.
%% Shortcoming 1: quality
First, the quality of metadata in the majority of participating archives remains
low.  According to the OLAC report card system, the average 
metadata-quality score for an OLAC record is 6.2 out of 10. However,
this score is
inflated by four of the larger archives that are already
following best practice; the average score for the remaining 31 archives
is only 3.5 out of 10.  Another way in which quality of current metadata suffers
is that it is not being kept up-to-date; most archives have not
updated their metadata to reflect new acquisitions since joining OLAC.
%% Shortcoming 2: quantity
Second, many language archives are not yet participating in OLAC.  For
example, the American Philosophical Association (Philadelphia) and the
National Anthropological Archives (Washington DC) are major national
centers having substantial collections of early language documentation
for native American languages.  The Oriental Institute (Chicago) and
the School of Oriental and African Studies (London) have vast
collections covering Asian and African languages.  Similarly, many
smaller archives are yet to participate in OLAC, e.g.\ the
Memorial University of Newfoundland Folklore and Language Archive (St John's);
the University of Wisconsin Himalayan Linguistics Archive (Milwaukee);
the College of William and Mary Creek Language Archive (Williamsburg).
Many linguistics departments have archives holding language
documentation collected over many decades by their staff, e.g. UCLA,
UC Santa Barbara, University of Chicago.
%% Shortcoming 3: OAIS
Third, many of the participating ``archives'' are more accurately described
as digitization projects.  Once project funding and institutional
backing cease, the materials may disappear.  Thus, these
initiatives are not yet following best practices in digital archiving,
e.g.\ as defined by the Open Archival Information Systems (OAIS)
reference model \citep{OAIS02} which has attained the status of an ISO
standard (ISO 14721:2003) and is accepted as best practice within the
digital library community.

%%%%%%%% Our vision for what needs to be done, long term...

In order to address these three shortcomings, we believe that the
language archives community needs to develop in three ways:

\renewcommand{\labelenumi}{T-1.\arabic{enumi}}
\begin{enumerate}\setlength{\itemsep}{0pt}
\item All OLAC repositories should have up-to-date catalogs
      that contain metadata conforming to best practice.
\item All major language archives should be participating in OLAC.
\item All OLAC repositories should conform to current best practices
      for the long-term curation of their holdings.
\end{enumerate}

\subsection{Access to Language Resources on the Web}

%%%%%%%% The status quo...

Language resources of all kinds have been posted on the web and can be
discovered with a standard search engine: interlinear texts,
dictionaries, linguistic descriptions, and linguistic research papers.
Even ordinary web pages may count as primary language documentation
when they are in any of the thousands of ``low-density'' languages,
languages with a small presence on the web.  Another source of
language resources is traditional libraries with their collections of
printed dictionaries, grammars, and linguistic monographs.  Library
automation has resulted in standards like MARC for digital cataloging
\citep{MARC}
and the Z39.50 protocol for searching digital catalogs \citep{Z3950}.
Today such catalogs typically have a web interface, though their entries are
usually not indexed by search engines.
WorldCat\footnote{\scriptsize\url{http://www.oclc.org/worldcat/}} is a single service
that provides a web-based search engine over the world's major
libraries. A more recent trend in the library automation arena has
been the development of e-print repositories, often in connection with
an academic institution.  These use simpler standards, Dublin Core for
metadata cataloging and the OAI protocol for
metadata harvesting.  OAIster is a single service that provides a
web-based search engine over all known e-print repositories.

%%%%%%%%

The task of finding language resources on the web is beset with
problems.  The most obvious one is \emph{scale}: that is, searching for
low-density language resources among the billions of pages indexed
by Google is like searching for a needle in a haystack.  Some
linguistic web-mining projects are
addressing this problem by probing search engines with linguistic
query terms and archiving the found objects
\citep{Langendoen02,Lewis03,BaldwinBird06}.  Another problem is \emph{volatility}:
for instance, a four-year longitudinal study has shown the half-life
of a web page to be approximately two years \citep{Koehler02}.
Yet another problem is that of the
\emph{hidden web}: namely, the content that lies behind search
interfaces and is thus opaque to conventional search engines.
Within the library automation community, JZKit\footnote{\scriptsize\url{http://developer.k-int.com/jzkit2/}} is an open-source project
that addresses this by building an aggregated index of library
catalogs that are accessible through about a dozen search interfaces
(including the Z39.50 and SRW/SRU\footnote{\scriptsize\url{http://www.loc.gov/standards/sru/}} standards).

%%%%%%%% Three shortcomings...

In order for the language resources community to realize the promise of universal access to relevant resources, three major
problems still need to be addressed.
%% Shortcoming 1: low precision and recall with Google
First, due to the huge scale of the search space and the absence
of precise indexing vocabularies,
users of web search engines such as Google typically experience
low precision and recall when searching for language resources.
Searches for scarce resources are often swamped with irrelevant
results (low precision).  For example searches for ``Santa Cruz'' will
not yield results for the East Papuan language of this name, spoken in
the Solomon Islands.
%%
Furthermore, many resources are just not returned at all because
search terms do not match the synonymous terms used in the desired
documents (low recall).  For example, searching for ``Dschang lexicon''
will not return the Dschang lexicon developed by one of the PIs
\citep{BirdTadadjeu97}; it was published as ``Lexique Y\'emba,'' since
the language of wider communication is French, and the language
autonym for \textit{Dschang} is \textit{Y\'emba}.
%%
Such materials are straightforward to find, however, using
linguistically-aware web mining programs that probe
Google's index by querying for all synonyms and translations of
linguistic terminology and by searching for pages that contain IPA characters.
A harder problem is to reliably identify the subject language and the
linguistic type of the found resource \citep{HughesBaldwin06lrec}.

%% Shortcoming 2: hidden web
A second problem is that the library automation solutions are part of
the hidden web and remain hidden to the language resources community
at large.
Users searching for language
resources also need to visit other services like WorldCat and OAIster;\footnote{\scriptsize\url{http://oaister.umdl.umich.edu/}}
it would be better for the language-resource content
of these services to be fully integrated with OLAC.
For example, OAI e-print repositories include many articles
relevant to languages and linguistics,\footnote{e.g.\ \small\url{http://romeo.eprints.org/search.php?t=Linguistics}}
for which results are not currently returned in OLAC searches.
Similarly, searching the world's libraries could be done with a
Google-mediated search of the WorldCat site, but this would not
permit systematic aggregation and integration, and would still suffer
from the problems of low recall and precision.
%% Shortcoming 3: finding OLAC
A third shortcoming is that users who try to find language resources
using any of these non-OLAC services are unlikely to discover that
OLAC can provide additional value, such as richer metadata and more
focused result sets.

%%%%%%%% Our vision for what needs to be done, long term...

In order to address these three shortcomings, we believe that the
language archives community needs to develop in three ways:

\renewcommand{\labelenumi}{T-2.\arabic{enumi}}
\begin{enumerate}\setlength{\itemsep}{0pt}
\item Low-density language materials identified in linguistic web-mining projects
      should be reliably categorized with OLAC vocabularies.
\item All OAI and library holdings relevant to language
      documentation should be indexed in OLAC, by
      crosswalking and enriching OAI-DC and MARC records.
\item Web search engines should index all OLAC records, so that
      users who discover language resources using a conventional 
      web search quickly find OLAC records and are drawn to the
      OLAC site for more precise searching.
\end{enumerate}

\subsection{Access to Language Resources from the Field}

Multimedia and web technologies now deliver unprecedented access to
rich digital resources.  However, most searches for archived language
resources still do not lead to an online version of the resource.  In
many cases, the resource is not digital, so that it is necessary to
visit an archive to access the physical artifact.  In other cases, the
resource is digital but the archive has never implemented a system for
automated access and it is necessary to contact an archivist or an
author to get a copy.  Even worse, there is a large quantity of
language documentation material that is still in the possession of the linguists who
originally collected it and has never been archived or cataloged.
As long as they remain unarchived, they are in imminent danger of
being lost to posterity \citep{Simons06}.  As long as these 
resources remain uncataloged, searches for language resources 
will never discover them.

%%%%%%%% DSpace repository as solution

What would help the language resource community to solve these problems is
for language archives to operate digital archiving systems that
automate the basic functions of curating a collection.  The Open
Archival Information Systems (OAIS) reference model is
an ISO standard that specifies best practices for
digital archiving \citep{OAIS02}.  It defines standards for the six basic functions
that any archive must perform: ingest, data management, archival
storage, access, preservation planning, and administration. and is
accepted as best practice within the digital library community.

In recent years, DSpace has emerged as the leading implementation of
a system that supports the OAIS reference model.  It also supports
the OAI protocol for metadata harvesting so that materials deposited
in a DSpace repository automatically become part of the virtual
digital library as indexed by OAIster and other OAI service providers.
DSpace was developed 
jointly in 2002 by the MIT Libraries and Hewlett-Packard Company to be
``MIT's online institutional repository---built to save, share, and
search MIT's digital research materials.''\footnote{\scriptsize\url{http://libraries.mit.edu/dspace-mit/}}
From the outset DSpace has been distributed as an open source software
product at
SourceForge\footnote{\scriptsize\url{http://sourceforge.net/projects/dspace/}}
where it is supported by a thriving developer community.  It is now in use at
more than one hundred universities as the institutional repository
where faculty members log in to deposit the fruits of their
research.\footnote{\scriptsize\url{http://wiki.dspace.org/DspaceInstances}}
The same software automatically supports search and access by the public,
including the ability to limit access where public access is not 
appropriate.  Given the large uptake by
major academic institutions along with the active open source mode of
development and dissemination, DSpace seems destined to be the leading
digital repository system for years to come. Furthermore, the future of
stored materials is ensured beyond the life of DSpace since the system
supports portable export formats for the archived information.

%%%%%%%% Three shortcomings...

Most language archives have not yet implemented automated ingest and
access for digital materials.  A system like DSpace would provide 
an off-the-shelf solution.  
However, before DSpace can meet the needs of digital language archives,
three major problems will need to be addressed.
% 1. precise language identification is missing
The first problem is the need to support precise language identification.
Language identification is the single most important thing that
characterizes a language resource; this is especially critical for
endangered languages---each relevant resource is highly valuable but
without precise language identification it becomes another needle in
the haystack. DSpace supports the ISO 639-2 standard for coding the
language that a resource is in \citep{ISO639-2}. Two changes are needed:
the ability to identify the language a resource is \emph{about}, and
the ability to use the ISO 639-3 standard.  ISO 639-2 provides
three-letter codes for fewer than 400 individual languages and covers all
remaining languages with 68 collective codes like [aus] ``Australian 
languages'' or [sai] ``South American Indian (other).''  This clearly
is not acceptable for indexing low-density language resources.
By contrast, the newly developed ISO 639-3 (which is being put to
the vote for final adoption this year) provides approximately 7,500 
three-letter codes for uniquely identifying all known human languages,
past and present \citep{ISO639-3}.  ISO 639-3 has already been adopted by OLAC as its
standard for language identification.
Adding support for ISO 639-3 to DSpace would not only meet the
needs of language resource creators, but would also have the
effect of getting precise language identification into the 
digital archiving mainstream, so that catalogers in general could
give precise identification that would already be in place
when we crosswalk from DC to OLAC.

% management of codespace over time deferred for section on method

% 2. DSpace is only for documents
A second problem is that the default configuration of DSpace
is for depositing conventional documents, such as books and articles.
However, language documentation involves other formats like
XML files, audio files, image files, and video files. Furthermore,
language resources need a richer set of metadata using the OLAC 
metadata standard with its specialized vocabularies for indexing.

% 3. fieldwork mode requires offline solution
A third problem to be solved is that linguists working
on language documentation in the field typically do not have internet access from
the fieldwork site. Thus they are unable to submit their
results to an institutional repository from the point of creation.
However, it is important to capture all the needed metadata at the time
of  creation.  Given fieldwork's ever-present risk of accident (such as
canoe tipping) or calamity (such as theft) or natural disaster (such
as tsunami) or operator error (such as inadvertent deletion), it is
also important to to be able to upload
collected materials to the institutional repository for safekeeping
as soon as it is possible to reach a location with internet access.
Therefore, the field linguist requires a solution in which archival
submissions can be created and fully documented while offline and then
uploaded later to the institutional repository.

%%%%%%%% Our vision for what needs to be done, long term...

In order to address these three shortcomings, we believe that the
language archives community needs to develop in three ways:

\renewcommand{\labelenumi}{T-3.\arabic{enumi}}
\begin{enumerate}\setlength{\itemsep}{0pt}
\item Users of an institutional repository should be able
  to specify language identification with precision.

\item Users of an institutional repository should be
    able to select an OLAC module so that they can integrate OLAC
    metadata creation into the workflow and lifecycle of their
    digital objects.

\item Field linguists should be able to participate in all the
    above while they are in the field.
\end{enumerate}


