\documentclass[11pt]{article}
\usepackage{times,url}

\addtolength{\textwidth}{1.6in}
\addtolength{\textheight}{1.6in}
\setlength{\hoffset}{-0.8in}
\setlength{\voffset}{-1in}
\setlength{\parskip}{1ex}
\setlength{\parindent}{0ex}

\renewcommand{\thesection}{\Alph{section}}

\newenvironment{reflist}{\begin{list}{}{\leftmargin 1em \itemindent -1em
    \listparindent \itemindent \parsep 0em \itemsep 0em}}{\end{list}}
\textheight 23.0cm

\begin{document}
\begin{center}
{\large\bf Biographical Sketch\\[15pt] Steven Bird}
\end{center}

\section{Professional Preparation}

\begin{tabular}{lll}
University of Melbourne, Australia & Computer Science and Mathematics & B.Sc. (hons) 1985 \\
University of Melbourne, Australia & Computer Science &  M.Sc. 1987 \\
University of Edinburgh, UK & Cognitive Science & Ph.D. 1991 \\
\end{tabular}

\section{Appointments}

\begin{enumerate}
\item Associate Professor, Department of Computer Science and
  Software Engineering, University of Melbourne (2002--present);

\item Senior Research Associate, Linguistic Data Consortium,
  University of Pennsylvania (2002--present);

\item Associate Director, Linguistic Data Consortium,
  Adjunct Associate Professor, Computer and Information Science, and
  Adjunct Associate Professor, Linguistics,
  University of Pennsylvania (1998--2002);

\item Research Fellow, Centre for Cognitive Science,
    University of Edinburgh (1990--98).
\end{enumerate}

\section{Publications}

\subsection{Five Relevant Publications}

{\raggedright
\begin{reflist}
\item Bird, S, Y Chen, S Davidson, H Lee, \& Y Zheng (2006).
  Designing and Evaluating an XPath Dialect for Linguistic Queries.
  {\it Proceedings of the 22nd International Conference on Data Engineering (ICDE)}
  pp 52--61.

\item Goldman, J, S Renals, S Bird, F de Jong, M Federico, C
  Fleischhauer, M Kornbluh, L Lamel, D Oard, C Stewart \& R Wright (2005).
  Accessing the Spoken Word, to
  appear in {\it International Journal on Digital Libraries} 5.

\item Bird, S \& G Simons (2004).
  Building an Open Language Archives Community on the DC Foundation.
  In Hillmann and Westbrooks (editors),
  \textit{Metadata in Practice: A Work in Progress}, ALA Editions, pp 203--222.

\item Simons, G \& S Bird (2003).
  Building an Open Language Archives Community on the OAI Foundation,
  to appear in {\it Library Hi Tech} 21, 210--218.
  Special Issue on the Open Archives Initiative.

\item Bird, S \& G Simons (2003).
  Seven Dimensions of Portability for Language Documentation and Description,
  {\it Language} 79, 557--582.
\end{reflist}}

\subsection{Five Other Publications}

{\raggedright
\begin{reflist}
\item Simons, G \& S Bird (2003).
  The Open Language Archives Community: An infrastructure for distributed archiving of language resources
  {\it Literary and Linguistic Computing} 18: 117-128.

\item Bird, S \& G Simons (2003).
  Extending Dublin Core Metadata to support the description and discovery of
  language resources, {\it Computing and the Humanities} 37, 375--388.

\item Bird, S \& M Liberman (2001).
  A formal framework for linguistic annotation.
  {\it Speech Communication}, 33, 23--60.

\item Bird, S \& G Simons (2000).
  {\it Web-Based Language Documentation and Description}.
  \url{http://www.ldc.upenn.edu/exploration/expl2000/}

\item Bird, S (1999). Multidimensional exploration of
  online linguistic field data.  {\it Proceedings of
  the 29th Annual Meeting of the Northeast Linguistics Society},
  pp 33--47.

\end{reflist}}

\section{Synergistic Activities}

\begin{reflist}
\item
  Coordinator of the Open Language Archives Community, a worldwide virtual
  library of language resources [\url{www.language-archives.org}]
\item
  Editor of ACL Digital Anthology [\url{acl.ldc.upenn.edu}]
\item Guest editor of a special issue
  of {\it Speech Communication} on speech annotation and corpus tools.
\end{reflist}

\section{Collaborators and Other Affiliations}

\begin{enumerate}
\item \textit{Collaborators/co-editors:}
  Catherine Bow (U Melbourne),
  Peter Buneman (U Edinburgh),
  Christopher Cieri (U Penn),
  Susan Davidson (U Penn),
  Franciska de Jong (U Twente),
  Marcello Federico (U Trento),
  Carl Fleischhauer (Library of Congress),
  Jerry Goldman (U Chicago),
  Jonathan Harrington (U Kiel, Germany),
  Chu-Ren Huang (Academia Sinica, Taiwan),
  Baden Hughes (U Melbourne),
  Ewan Klein (U Edinburgh),
  Mark Kornbluh (Michigan State),
  Lori Lamel (LIMSI),
  Haejoong Lee (U Penn),
  Mark Liberman (U Penn),
  Edward Loper (U Penn),
  Xiaoyi Ma (U Penn),
  Brian MacWhinney (CMU),
  Kazuaki Maeda (U Penn),
  Craig Martell (U Penn),
  Douglas Oard (U Maryland),
  Steve Renals (U Edinburgh),
  Gary Simons (SIL),
  Claire Stewart (Northwestern U),
  Richard Wright (BBC).
\item \textit{Graduate Advisors:} 
  Roland Sussex, University of Queensland, Australia;
  Ewan Klein, University of Edinburgh, UK;
  Robin Cooper, G\"oteborg University, Sweden.
\item \textit{recent advisees or thesis committee connections:}
  John Bell,
  Phil Blunsom,
  Trevor Cohn,
  Rod Farmer,
  Catherine Lai,
  Patrick Ye.
\item \textit{Total number of Ph.D. students:} 10
\end{enumerate}

\end{document}
