\documentclass[11pt]{article}
\usepackage{times,url}

\addtolength{\textwidth}{1.6in}
\addtolength{\textheight}{2.0in}
\setlength{\hoffset}{-0.8in}
\setlength{\voffset}{-1in}
\setlength{\parskip}{1ex}
\setlength{\parindent}{0ex}

\renewcommand{\thesection}{\Alph{section}}

\newenvironment{reflist}{\begin{list}{}{\leftmargin 1em \itemindent -1em
    \listparindent \itemindent \parsep 0em \itemsep 0em}}{\end{list}}
\textheight 23.0cm

\begin{document}
\begin{center}
{\Large\bf Biographical Sketch\\[15pt] Gary F.~Simons}
\end{center}

\section{Professional Preparation}

\begin{tabular}{lll}
Seattle Pacific University & Interlanguage & B.A. (Summa Cum Laude) 1974 \\
Cornell University & Linguistics & M.A. 1976 \\
Cornell University & Linguistics & Ph.D. 1979 \\
\end{tabular}

\section{Appointments}

\begin{reflist}
\item Associate VP for Academic Affairs, SIL International (1999--)

\item Adjunct Associate Professor of Language Development,
  Graduate Institute of Applied Linguistics (1999--)

\item Adjunct Assistant Professor of Linguistics,
  University of Texas at Arlington (1985--)

\item Director, Academic Computing, SIL (1986--1999)

\item Manager, Language Data Processing, SIL (1984--1985)

\item Translation Advisor for North Malaita, Solomon Islands (with SIL, 1979--1983)

\item Graduate Research Assistant, Cornell University (1976--1978)

\end{reflist}

\section{Publications}

\subsection{Five Relevant Publications}

{\raggedright
\begin{reflist}
\item Bird, Steven and Gary Simons (2004).
  Building an Open Language Archives Community on the DC Foundation.
  In Diane Hillmann and Elaine Westbrooks (eds.),
  {\it Metadata in Practice,} pp. 203--222. Chicago: American Library Association.
  Preprint: \url{http://www.sil.org/~simonsg/preprint/Metadata%20in%20Practice.pdf}

\item Bird, Steven and Gary Simons (2003).
  Seven Dimensions of Portability for Language Documentation and Description,
  {\it Language} 79(3):557--582.
  Preprint:~\url{http://www.sil.org/~simonsg/preprint/Seven%20dimensions.pdf}

\item Simons, Gary and Steven Bird (2003).
  Building an Open Language Archives Community on the OAI Foundation,
  {\it Library Hi Tech} 21(2):210--218.
  Special Issue on the Open Archives Initiative. 
  Preprint:~\url{http://arxiv.org/abs/cs.CL/0302021}

\item Simons, Gary and Steven Bird (2003).
  The Open Language Archives Community: An infrastructure for distributed archiving of language resources
  {\it Literary and Linguistic Computing} 18(2):117-128.
  Preprint:~\url{http://arxiv.org/abs/cs.CL/0306040}

\item Bird, Steven and Gary Simons (2003).
  Extending Dublin Core Metadata to support the description and discovery of
  language resources, {\it Computers and the Humanities} 37(4):375--388.
  Preprint:~\url{http://arxiv.org/abs/cs.CL/0308022}

\end{reflist}}

\subsection{Five Other Publications}

{\raggedright
\begin{reflist}
\item Simons, Gary F. (2006).
  Ensuring that digital data last: The priority of archival form over working form and presentation form.
  {\it SIL Electronic Working Papers} 2006-003. Dallas: SIL International.
  Online:~\url{http://www.sil.org/silewp/abstract.asp?ref=2006-003}


\item Simons, Gary F., William D. Lewis, Scott O. Farrar, D. Terence Langendoen, Brian Fitzsimons, and Hector Gonzalez (2004).
  The semantics of markup: Mapping legacy markup schemas to a common semantics.
  In Graham Wilcock, Nancy Ide, and Laurent Romary (eds.),
  {\it Proceedings of the 4th workshop on NLP and XML  (NLPXML-2004),} pp. 25--32.
  Association for Computational Linguistics.
  Preprint:~\url{http://linguistlist.org/emeld/emeld/documents/index.cfm}


\item Simons, Gary F. (1998).
  The nature of linguistic data and the requirements of a computing environment for linguistic research.
  In John Lawler and Helen Aristar Dry (eds.), {\it Using Computers in Linguistics: A Practical Guide,} pp. 10-–25.
  London and New York: Routledge. 

\item Pike, Kenneth L. and Gary F. Simons (1996).
  Toward the historical reconstruction of matrix patterns in morphology.
  In Kenneth L. Pike, Gary F. Simons, Carol V. McKinney, and Donald Burquest, 
  {\it The Mystery of Cultural Contacts, Historical Reconstruction, and Text Analysis,} pp. 1-–37.
  Georgetown University Press.

\item Langendoen, D. Terence and Gary F. Simons (1995).
  A rationale for the TEI recommendations for feature-structure markup,
  {\it Computers and the Humanities} 29(3):191–-209.
  [Reprinted in Nancy Ide and Jean Veronis (eds.),
  {\it The Text Encoding Initiative: Background and Context,} pp. 191-–209.
  Dordrecht: Kluwer Academic Publishers.]


\end{reflist}}

\section{Synergistic Activities}

\begin{reflist}
\item
  Coordinator (with Steven Bird) of the Open Language Archives Community, a worldwide virtual
  library of language resources [\url{www.language-archives.org}] (2000--)
\item
  Developer (with Peter Constable) of {\it ISO 639-3: Codes for the representation of names of languages}
  (providing unique three-letter identifiers for approximately 7,500 human languages, past and present)
  and developer of the Registration Authority web site [\url{www.sil.org/iso639-3/}] (2002--)
\item
  Executive editor of {\it Ethnologue: Languages of the World}, print and web versions, fifteenth edition (published 2005)
  and sixteenth edition (in preparation)
\item
  External consultant on past NSF-sponsored projects
  for Linguist List (Eastern Michigan U)---Software Development for the {\small LINGUIST} Network,
  The {\small LINGUIST} Multi-List Support Project, Database Design for Endangered Languages Data,
  E-NELD: Electronic Metastructures for Endangered Language Data---and for the 
  Linguisitic Data Consortium (U Pennsylvania)---Linguistic Exploration,
  International Standards for Language Engineering, TalkBank.
  Consultant to three current projects---LL-Map: Language and Location (Eastern Michigan U),
  Multi-Tree: A Digital Library of Language Relationships (Wayne State U), and 
  Five Languages of Eurasia: Field Work, Analysis and Digital Archiving (Colgate U)---with combined commitment of 12 days per year.

\item
  Committee on Endangered Languages and their Preservation,
  Linguistic Society of America (2006--)

\end{reflist}

\section{Collaborators and Other Affiliations}

\begin{enumerate}
\item \textit{Collaborators/co-eds.:}
  Anthony Aristar (Wayne State U),
  Steven Bird (U Melbourne),
  Peter Constable (Microsoft),
  Helen Aristar Dry (Eastern Michigan U),
  Scott Farrar  (U Washington), 
  Hector Gonzalez (California State U, Fresno),
  Larry Hayashi (SIL International),
  Chu-Ren Huang (Academia Sinica, Taiwan),
  Baden Hughes (U Melbourne),
  D. Terence Langendoen (NSF),
  William Lewis (U Washington), 
  Mike Maxwell (CASL),
  Alexander Nakhimovsky (Colgate U)

\item \textit{Graduate Advisor:} 
  Joseph Grimes (SIL and U Hawaii at Manoa)

\item \textit{Thesis advisees:}
  Eric Albright (SIL),
  Eugene Gruber (?),
  Linda Humnick (SIL),
  Lars Huttar (SIL)
  Taeho Jang (SIL),
  Ken Prettol (SIL),
  John Wimbish (SIL). 
  \it{Total:} 7 

\end{enumerate}
\end{document}
